---
title: "Ch10"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(nycflights13)
```

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

You would need to combine `airports` with `flights` because the `airports` dataset the the coordinates of the airport. You could match them by the `faa` variable in `airports` and the `origin` and `dest` from `flights`

```{r}
flights %>%
  left_join(airports, by = c("origin" = "faa", "dest" = "faa"))
```

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

Similarly as the above exercise, you can match them with `faa` in `airports` and `origin` in `weather`.

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

You could also connect it with flights through the `dest` and have the weather of every single airport in the U.S that are present in the `flights` dataset.

We know that some days of the year are "special"", and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

You could have a separate dataset with the festivities in the U.S by day and month. With this information you can match it with each flight in the `flights` data set and subsequently with the `weather` dataset

## Exercises 13.3.1

Add a surrogate key to flights.

```{r}
flights %>%
  mutate(id = row_number(year)) %>%
  select(id, everything())
```

Identify the keys in the following datasets

Lahman::Batting,

```{r}
as_tibble(Lahman::Batting)
```

It's `playerID`.

babynames::babynames

```{r}
as_tibble(babynames::babynames)
```

I think it might be the combination of `year` and `name`. That identifies each name-year pair for matching with other tables.

nasaweather::atmos

```{r}
as_tibble(nasaweather::atmos)
```

Here it's most likely `lat`, `long`, `year` and `month`, which locate a specific place in a month/year.

fueleconomy::vehicles

```{r}
as_tibble(fueleconomy::vehicles)
```
`id` is the simple key.

ggplot2::diamonds

```{r}
as_tibble(ggplot2::diamonds)
```

There is not key because there are not other datasets! The concept of key only makes sense when there are other relational datasets.

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

It's actualy *very* straight forward: all three tables have the same `playerID` and `yearID` and each table has the information that the other doesn't have, so they complement each other. I think it is one-to-one relationships but that needs to have inspected further.

## Exercises 13.4.6

Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:


```{r}
flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  group_by(dest) %>%
  summarize(avg_delay = mean(tot_delay, na.rm = TRUE)) %>%
  left_join(select(airports, faa, lon, lat), c("dest" = "faa")) %>%
  ggplot(aes(lon, lat, colour = avg_delay)) +
  borders("state") +
  geom_point(size = 2, alpha = 0.8) +
  xlim(c(-130, -65)) +
  ylim(c(20, 50)) +
  coord_quickmap() +
  viridis::scale_color_viridis()
```


Add the location of the origin and destination (i.e. the lat and lon) to flights.

```{r}
flights %>%
  left_join(select(airports, faa, lat, lon), by = c("origin" = "faa")) %>%
  rename(lat_origin = lat,
         lon_origin = lon) %>%
  left_join(select(airports, faa, lat, lon), by = c("dest" = "faa")) %>%
  rename(lat_dest = lat,
         lon_dest = lon) %>%
  select(origin, dest, matches("lat|lon"))
```

(However, this could've been done with `gather` and avoid the double `left_join`)

Is there a relationship between the age of a plane and its delays?

```{r}
flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  group_by(tailnum) %>%
  summarize(avg_delay = mean(tot_delay, na.rm = TRUE)) %>%
  left_join(select(planes, tailnum, year), by = "tailnum") %>%
  mutate(year = 2013 - year) %>%
  ggplot(aes(avg_delay, year)) +
  geom_point() +
  geom_smooth()
```

From a very preliminary view, there doesn't seem to be, although the some more older planes have very short delays and some younger planes have very high delays. This pattern however could be due to other things such as the origin/destionation.

What weather conditions make it more likely to see a delay?

```{r}
avg_del <-
  flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  group_by(month, day) %>%
  summarize(avg_delay = mean(tot_delay, na.rm = TRUE))

avg_weather <-
  weather %>%
  group_by(month, day) %>%
  select(-hour) %>%
  summarize_at(vars(temp, humid, wind_speed, precip), mean, na.rm = TRUE)

avg_del %>%
  left_join(avg_weather) %>%
  ungroup() %>%
  mutate(avg_delay = cut_width(avg_delay, 35)) %>%
  gather(weather, metrics, -(month:avg_delay)) %>%
  ggplot(aes(avg_delay, metrics)) +
  geom_boxplot() +
  facet_wrap(~ weather, scales = "free_y")
```

Humidity seems to be the one more related to delays, although this is in isolation. A more thorough approach would be to create a grid of all possible combinations of weather conditions and match them up with the delay and then compare the combination of weather conditions with the highest delays.

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

Here is for Junte 13 2013:
```{r}
flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  group_by(month, day, dest) %>%
  summarize(avg_delay = mean(tot_delay, na.rm = TRUE)) %>%
  filter(month == 6, day == 13) %>%
  left_join(select(airports, faa, lat, lon), by = c("dest" = "faa")) %>%
  ggplot(aes(lon, lat, colour = avg_delay)) +
  borders("state") +
  geom_point(size = 2, alpha = 0.8) +
  xlim(c(-130, -65)) +
  ylim(c(20, 50)) +
  coord_quickmap() +
  viridis::scale_color_viridis()
```

I didn't know if that was a big increase in delays so I cross-checked randomly with the next month:

```{r}
flights %>%
  mutate(tot_delay = arr_delay + dep_delay) %>%
  group_by(month, day, dest) %>%
  summarize(avg_delay = mean(tot_delay, na.rm = TRUE)) %>%
  filter(month == 7, day == 13) %>%
  left_join(select(airports, faa, lat, lon), by = c("dest" = "faa")) %>%
  ggplot(aes(lon, lat, colour = avg_delay)) +
  borders("state") +
  geom_point(size = 2, alpha = 0.8) +
  xlim(c(-130, -65)) +
  ylim(c(20, 50)) +
  coord_quickmap() +
  viridis::scale_color_viridis()
```

You can see big differences in the east coast where in June the average delay was much higher. Apparently there was a sever weather storm in the east coast. See [here](https://www.washingtonpost.com/news/capital-weather-gang/wp/2013/06/14/june-13-2013-severe-weather-hypestorm-or-the-real-deal/?utm_term=.ec80520fb97c)

## Exercises 13.5.1

What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
flights %>%
    anti_join(planes, by = "tailnum") %>%
    count(carrier, sort = TRUE)
```

AA and MQ don't seem to report tail numbers.

Filter flights to only show flights with planes that have flown at least 100 flights.

```{r}
flights %>%
  semi_join(count(flights, tailnum) %>% filter(n >= 100))
```


Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

```{r}
ten_common <-
  fueleconomy::common %>%
  ungroup() %>%
  arrange(-n) %>%
  top_n(10, n)

fueleconomy::vehicles %>%
  semi_join(ten_common)
```


Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

```{r}
fn <-
  flights %>%
  group_by(month, day) %>%
  summarize(avg_delay = sum(arr_delay + dep_delay, na.rm = TRUE)) %>%
  mutate(twoday_delay = avg_delay + lag(avg_delay)) %>%
  arrange(-twoday_delay)

wea <-
  weather %>%
  group_by(month, day) %>%
  summarize_at(vars(humid, precip, temp), mean, na.rm = TRUE)

fn %>%
  left_join(wea) %>%
  arrange(twoday_delay)
```

Very superficially, it seems as though precipitation was higher in delays flights, and temprature was slightly higher. Although the pattern is just by looking at top/bottom 10. Graphical inspection should yield more interesting patterns.


What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?

```{r}
anti_join(flights, airports, by = c("dest" = "faa"))
# Give me the flights from the destionations that that are not present in the `airports` dataset.


anti_join(airports, flights, by = c("faa" = "dest"))
# Give me the airports that are not present as destinations in the `flights` dataset.
```


You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

```{r}
flights %>%
  group_by(tailnum, carrier) %>%
  count() %>%
  filter(n() > 1) %>%
  select(tailnum) %>%
  distinct(tailnum)
  

```

