---
title: "Ch10"
output: html_notebook
---

10.5 Exercises

```{r, echo = F, warnings = F}
library(tibble)
```

1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

```{r}
print(mtcars)
```

It's not simply because it doesn't have the right format.

```{r}
class(mtcars)
```

2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df_tibble <- tibble::tibble(abc = 1, xyz = "a")

# df does partial matching
df$x
df_tibble$x # tibble doesn't
```

```{r}
df[, "xyz"] # returns a factor
df_tibble[, "xyz"] # returns a data frame
```

```{r}
# returns the same thing but xyz is a factor in the data farme and a char in the tibble
df[, c("abc", "xyz")]
df_tibble[, c("abc", "xyz")]
```

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r, eval = F}
var <- "abc"

# Will extract the vector
df_tibble[[var]]

# Will select the var and return a tibble
df_tibble[var]
```

4. Practice referring to non-syntactic names in the following data frame by:
```{r}
annoying <- tibble::tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

Extracting the variable called 1.
```{r}
annoying$`1`
```

Plotting a scatterplot of 1 vs 2.
```{r}
ggplot(annoying, aes(x = `1`, y = `2`)) + geom_point()
```

Creating a new column called 3 which is 2 divided by 1.
```{r}
annoying <-
  annoying %>%
  mutate(`3` = `2`/`1`)
```

Renaming the columns to one, two and three.
```{r}
annoying %>%
  rename(one = `1`,
         two = `2`,
         three = `3`)
```

5. What does tibble::enframe() do? When might you use it?

It turns a vector into a tibble. If the vector is named the it assigns the name as the column name.

```{r}
x <- letters[1:10]
enframe(x)

m <- enframe(list(a = 5, b = 7))
```

One application is when having a named list that you'd like to turn into a dataframe. That list might contain regression models or data sets, for example.

What option controls how many additional column names are printed at the footer of a tibble?
tibble.max_extra_cols

```{r}
library(fivethirtyeight)
options(tibble.max_extra_cols = 5)

police_killings
```
