---
title: "Ch10"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(nycflights13)
```

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

You would need to combine `airports` with `flights` because the `airports` dataset the the coordinates of the airport. You could match them by the `faa` variable in `airports` and the `origin` and `dest` from `flights`

```{r}
flights %>%
  left_join(airports, by = c("origin" = "faa", "dest" = "faa"))
```

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

Similarly as the above exercise, you can match them with `faa` in `airports` and `origin` in `weather`.

weather only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

You could also connect it with flights through the `dest` and have the weather of every single airport in the U.S that are present in the `flights` dataset.

We know that some days of the year are "special"", and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

You could have a separate dataset with the festivities in the U.S by day and month. With this information you can match it with each flight in the `flights` data set and subsequently with the `weather` dataset

## Exercises 13.3.1

Add a surrogate key to flights.

```{r}
flights %>%
  mutate(id = row_number(year)) %>%
  select(id, everything())
```

Identify the keys in the following datasets

Lahman::Batting,

```{r}
as_tibble(Lahman::Batting)
```

It's `playerID`.

babynames::babynames

```{r}
as_tibble(babynames::babynames)
```

I think it might be the combination of `year` and `name`. That identifies each name-year pair for matching with other tables.

nasaweather::atmos

```{r}
as_tibble(nasaweather::atmos)
```

Here it's most likely `lat`, `long`, `year` and `month`, which locate a specific place in a month/year.

fueleconomy::vehicles

```{r}
as_tibble(fueleconomy::vehicles)
```
`id` is the simple key.

ggplot2::diamonds

```{r}
as_tibble(ggplot2::diamonds)
```

There is not key because there are not other datasets! The concept of key only makes sense when there are other relational datasets.

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

It's actualy *very* straight forward: all three tables have the same `playerID` and `yearID` and each table has the information that the other doesn't have, so they complement each other. I think it is one-to-one relationships but that needs to have inspected further.