---
title: "Ch12"
output: html_notebook
---

12.2.1 Exercises

1. Using prose, describe how the variables and observations are organised in each of the sample tables.

table1:
This is a tidy data set. Country is a column containing all country names, year contains each year in stacked format and the remaining two variables are independent values of different measures.

table2:
Exactly as the previous table, but the type and count are actually storing two variables into one. Both variables have a different metric and thus should be different variables.

table3:
Has some tidy principles but as the previous table, it combines two columns in one in the last column.

table4a and table4b:

Both tables are untidy because the year variable should be stacked.

2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

Extract the number of TB cases per country per year.

```{r}
devtools::install_github("garrettgman/DSR")
library(DSR)

tidy_tb <-
  table2 %>%
  spread(type, count)

cases <-
  tidy_tb %>%
  select(-population)
```

Extract the matching population per country per year.

```{r}
population <-
  tidy_tb %>%
  select(-cases)
```
Divide cases by population, and multiply by 10000.

Store back in the appropriate place.

```{r}
tidy_tb <- 
  tidy_tb %>%
  mutate(rate = cases$cases/population$population)
```

Which representation is easiest to work with? Which is hardest? Why?

The first representation. I could've computed everything in one pipeline simply because the data was stacked rather than wide. Doing column operations is extremely easy in wide format, so `spread` is particularly useful for transformations and then turning back.

3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

```{r}
table2 %>%
  spread(type, count) %>%
  mutate(year = as.numeric(year)) %>%
  ggplot(aes(year, cases)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))
```

12.3.3 Exercises

1. Why are gather() and spread() not perfectly symmetrical?
Carefully consider the following example:

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)

stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
# (Hint: look at the variable types and think about column names.)
```
Because in most cases when you `gather` a variable the stacked column will be a categorical variable, thus gather turns it into a a character vector.

2. Both spread() and gather() have a convert argument. What does it do?

To turn a character vector into a logical, interger, numeric, of factor if appropriate.

3. Why does this code fail?

```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n
```
Because you need to `` the non-synthetic names.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```

4. Why does spreading this tibble fail? How could you add a new column to fix the problem?

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people %>%
  spread(key, value)
```
Because Phillip Woods has two ages, so no unique identifier. We could add a new column that uniquely identifies each person.

```{r}
people %>%
  mutate(id = c(1, 1, 2, 3, 3)) %>%
  spread(key, value)
```

5. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

```{r}

preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

preg %>%
  gather(gender, freq, -pregnant)
```

We gather it and turn the gender variable into a stacked column.

2.4.3 Exercises

What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

# In this case, the second row will have a missing. Should
# we want the first or second letters to be in the 2nd or 3rd column, for example?

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))

# By specifying fill we can fix that.
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")

# If the tibble would be like the first one:
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "drop")

# Then 'extra' controls what will happen with the additional letter

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")


```

Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

`remove` will remove the columns you want to turn into one single column in `unite` and will remove the pasted column to separate in `separate`.

Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

This one I still not find the differences

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  extract(x, c("one", "two", "three"))
```

```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  extract(x, c("one", "two"))
```


# 12.5.1 Exercises
Missing