---
title: "Ch12"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```

## Exercise 15.3.1
Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

Well, we should recode the levels so that all non-income categories are at the end and the plot is set to `coord_flip` so that the labels can be read.

```{r}
gss_cat %>%
  ggplot(aes(rincome)) +
  geom_bar()

gss_cat %>%
  mutate(rincome =
           fct_relevel(rincome,
                       c("No answer", "Don't know", "Refused", "Not applicable"))) %>%
  ggplot(aes(rincome)) +
  geom_bar() +
  coord_flip()
```


What is the most common relig in this survey? What’s the most common partyid?

```{r}
gss_cat %>%
  count(relig) %>%
  arrange(-n)
```
```{r}
gss_cat %>%
  count(partyid) %>%
  arrange(-n)
```


Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

```{r}
gss_cat %>%
  count(relig, denom) %>%
  filter(denom == "No denomination")
```
```{r}

```



