---
title: "Ch12"
output: html_notebook
---

12.2.1 Exercises

1. Using prose, describe how the variables and observations are organised in each of the sample tables.

table1:
This is a tidy data set. Country is a column containing all country names, year contains each year in stacked format and the remaining two variables are independent values of different measures.

table2:
Exactly as the previous table, but the type and count are actually storing two variables into one. Both variables have a different metric and thus should be different variables.

table3:
Has some tidy principles but as the previous table, it combines two columns in one in the last column.

table4a and table4b:

Both tables are untidy because the year variable should be stacked.

2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

Extract the number of TB cases per country per year.

```{r}
devtools::install_github("garrettgman/DSR")
library(DSR)

tidy_tb <-
  table2 %>%
  spread(type, count)

cases <-
  tidy_tb %>%
  select(-population)
```

Extract the matching population per country per year.

```{r}
population <-
  tidy_tb %>%
  select(-cases)
```
Divide cases by population, and multiply by 10000.

Store back in the appropriate place.

```{r}
tidy_tb <- 
  tidy_tb %>%
  mutate(rate = cases$cases/population$population)
```

Which representation is easiest to work with? Which is hardest? Why?

The first representation. I could've computed everything in one pipeline simply because the data was stacked rather than wide. Doing column operations is extremely easy in wide format, so `spread` is particularly useful for transformations and then turning back.

3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?

```{r}
table2 %>%
  spread(type, count) %>%
  mutate(year = as.numeric(year)) %>%
  ggplot(aes(year, cases)) +
  geom_line(aes(group = country), colour = "grey50") +
  geom_point(aes(colour = country))

```