---
title: "Ch12"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```

## Exercise 15.3.1
Explore the distribution of rincome (reported income). What makes the default bar chart hard to understand? How could you improve the plot?

Well, we should recode the levels so that all non-income categories are at the end and the plot is set to `coord_flip` so that the labels can be read.

```{r}
gss_cat %>%
  ggplot(aes(rincome)) +
  geom_bar()

gss_cat %>%
  mutate(rincome =
           fct_relevel(rincome,
                       c("No answer", "Don't know", "Refused", "Not applicable"))) %>%
  ggplot(aes(rincome)) +
  geom_bar() +
  coord_flip()
```


What is the most common relig in this survey? What’s the most common partyid?

```{r}
gss_cat %>%
  count(relig) %>%
  arrange(-n)
```
```{r}
gss_cat %>%
  count(partyid) %>%
  arrange(-n)
```


Which relig does denom (denomination) apply to? How can you find out with a table? How can you find out with a visualisation?

```{r}
gss_cat %>%
  count(relig, denom) %>%
  filter(denom == "No denomination")
```

## Exercises 15.4.1
There are some suspiciously high numbers in tvhours. Is the mean a good summary?

```{r}
gss_cat %>%
  ggplot(aes(tvhours)) +
  geom_histogram() +
  geom_vline(xintercept = mean(gss_cat$tvhours, na.rm = TRUE), colour = "red") +
  geom_vline(xintercept = median(gss_cat$tvhours, na.rm = TRUE), colour = "blue")
```

Nope, there's a reasonable differnece between the mean and the median.

For each factor in gss_cat identify whether the order of the levels is arbitrary or principled.

```{r}
fct_gss <- gss_cat[sapply(gss_cat, is.factor)]

lapply(fct_gss, levels)
```

For all variables except rincome the levels are arbitrary. `rincome` is the only one which has a principled order.

Why did moving 'Not applicable' to the front of the levels move it to the bottom of the plot?

```{r}
gss_cat %>%
  mutate(rincome = rincome %>% fct_relevel("Not applicable")) %>%
  ggplot(aes(rincome)) +
  geom_bar()
```

The previous plot moves it the beginning by if the flip the coordinates, it looks like it's in the end.

```{r}
gss_cat %>%
  mutate(rincome = rincome %>% fct_relevel("Not applicable")) %>%
  ggplot(aes(rincome)) +
  geom_bar() +
  coord_flip()
```

But that's an illusion because of changing the coordinates. The plot is in the same order.

## Exercises 15.5.1
How have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r}

all_levels <- levels(gss_cat$partyid)

gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    Democract = c('Not str democrat', 'Strong democrat'),
    Republican = c('Strong republican', 'Not str republican'),
    Independent = c("Ind,near rep", "Independent", "Ind,near dem"),
    Others = c("No answer", "Don't know", "Other party")
  )) %>%
  count(year, partyid) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  ggplot(aes(year, perc, group = partyid, colour = partyid)) +
  geom_line() +
  theme_bw()
```

It looks like Independents are growing whereas both other groups are shrinking, with Republicans shrinking much faster.

How could you collapse rincome into a small set of categories?

A very quick but perhaps not so advisable way (because you might lose important information and substantive meaning) is to use `fct_lump`.

```{r}
gss_cat %>%
  mutate(rincome = fct_lump(rincome, n = 6)) %>%
  count(rincome)
```
or

```{r}
gss_cat %>%
  mutate(rincome =
           fct_collapse(
             rincome,
             `Unknown` = c("No answer", "Don't know", "Refused", "Not applicable"),
             `Lt $5000` = c("Lt $1000", str_c("$", c("1000", "3000", "4000"),
                                              " to ", c("2999", "3999", "4999"))),
             `$5000 to 10000` = str_c("$", c("5000", "6000", "7000", "8000"),
                                      " to ", c("5999", "6999", "7999", "9999"))
           )) %>%
  ggplot(aes(x = rincome)) +
  geom_bar() +
  coord_flip()
```

