---
title: "Ch15"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
library(tidyverse)
library(purrr)
```

## Exercises 20.3.5
Describe the difference between is.finite(x) and !is.infinite(x).

There is no difference! `is.finite()` tests whether a value is **not** `-Inf` or `Inf` and `is.infinite` tests whether it is `-Inf` or `Inf`. Logically, `!` negates the previous.

However... `is.infinite` `NA` and `NaN` to be non infinite! See the answer to ex 19.3.1 [here](https://jrnold.github.io/r4ds-exercise-solutions/vectors.html).

Read the source code for dplyr::near() (Hint: to see the source code, drop the ()). How does it work?

It calculates absolute difference up to a certain tolerance level. If the difference is 0.2 it calculates the difference and compares it to the tolerance level.

A logical vector can take 3 possible values. How many possible values can an integer vector take? How many possible values can a double take? Use google to do some research.

Taken from: https://jrnold.github.io/r4ds-exercise-solutions/vectors.html

The help for .Machine describes some of this:

As all current implementations of R use 32-bit integers and uses IEC 60559 floating-point (double precision) arithmetic,

Brainstorm at least four functions that allow you to convert a double to an integer. How do they differ? Be precise.

```{r}
x <- seq(-10, 10, by = 0.5)
as.integer(x) # turns to a literal integer
floor(x) # rounds down
`%/%`(x, 1) # it extracts the integer part of x

```
I can't think of a 4th

What functions from the readr package allow you to turn a string into logical, integer, and double vector?

```{r}
x <- c("TRUE", "FALSE")
parse_logical(x)
x <- c("1", "2")
parse_integer(x)
x <- c("1.1", "1.2")
parse_double(x)
```

## Exercises 20.4.6
What does mean(is.na(x)) tell you about a vector x? What about sum(!is.finite(x))?

`mean(is.na(x))` gives the proportion or percentage of missing values in that vector
`sum(!is.finite(x))` gives the total number of infinite objects in `x`

Carefully read the documentation of is.vector(). What does it actually test for? Why does is.atomic() not agree with the definition of atomic vectors above?

`is.atomic()` tests for objects that area "logical", "integer", "numeric" (synonym "double"), "complex", "character" and "raw". `is.vector()` on the other hand tests for any of the `atomic` modes but must have no attributes other than names.

A factor, for example, shows how they contradict.

```{r}
x <- factor()
is.vector(x)
is.atomic(x)
```


Compare and contrast setNames() with purrr::set_names().

The do the same thing but `purrr::set_names` but with stricter argument checking and cool argument to rename variable based on a function.

Create functions that take a vector as input and returns:

The last value. Should you use [ or [[?

`[[` because it wants the specific value. `x` might have other attributes such as names and it is asking only for one value.
```{r}
returner <- function(x) {
  x[[length(x)]]
}

returner(1:10)
```


The elements at even numbered positions.

```{r}
even_num <- function(x) {
  x[seq(2, length(x), by = 2)]
}
even_num(1:10)
```


Every element except the last value.

```{r}
except_last <- function(x) {
  x[-length(x)]
}

except_last(1:10)
```


Only even numbers (and no missing values).

```{r}
even_numbers <- function(x) {
  x[!is.na(x) & x %% 2 == 0]
}

even_numbers(sample(c(NA, 1:100), 99))
```


Why is x[-which(x > 0)] not the same as x[x <= 0]?

`x[-which(x > 0)]` subset the values of `x` which ar at `0` or below
`x[x <= 0]` subset the values of x which ar at `0` or below.

They are the same.

However, `which` will ignore `NA`'s and leave them as is and `<=` will turn any value that cannot be comparable to `NA` like `NaN`!

```{r}
x <- c(-5:5, Inf, -Inf, NaN, NA)

x[-which(x > 0)]
x[x <= 0]
```
Notice the difference?


What happens when you subset with a positive integer that’s bigger than the length of the vector? What happens when you subset with a name that doesn’t exist?

```{r}
x <- 1:10
x[11]
```
Returns an `NA`

```{r}
x <- c("a" = 1)
x["b"]
```
Both the name and element is `NA`.

## Exercises 20.5.4
Draw the following lists as nested sets:

`list(a, b, list(c, d), list(e, f))`
`list(list(list(list(list(list(a))))))`

C'mon!! No paper and pen.

What happens if you subset a tibble as if you’re subsetting a list? What are the key differences between a list and a tibble?

```{r}
x <- as_tibble(mtcars)
x[1] # gives back a tibble (same as lists)
x[[1]] # gives back a vector (same as lists)
# In fact, data frames and tibbles are lists!!
```

## Exercises 20.7.4

What does hms::hms(3600) return? How does it print? What primitive type is the augmented vector built on top of? What attributes does it use?

```{r}
x <- hms::hms(3600)
```
Because sixty seconds multiplied by sixty minutes is an hour (3600 seconds) it returns `1` hour.

```{r}
class(x) # augmented classes
typeof(x)# a double, in fact, it is 3600 seconds if we use as.numeric
```

It uses two classes and units attribute. It prints probable as `cat` but it is in fact represented as seconds behind the scenes. The printing in hours is just a nice format to see.

Try and make a tibble that has columns with different lengths. What happens?

The length of all columns must be the same or 1 (scalar), otherwise throws an error.
```{r}
tibble(a = 1:2, b = 1:3)
```

Based on the definition above, is it ok to have a list as a column of a tibble?

Yes, as long as it has the same length as other columns. But beware, because to 'open' that column all slots need to have the same class.