---
title: "Ch17"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(nycflights13)
```

## Exercises 21.2.1

Write for loops to:

```{r}
looper <- function(the_type, object, fun) {
  empty_vec = vector(the_type, length(object))

  for (every_column in seq_along(object)) {
    empty_vec[every_column] <- fun(object[[every_column]])  
  }
  
  empty_vec
}
```


Compute the mean of every column in mtcars.

```{r}
looper("numeric", mtcars, mean)
```


Determine the type of each column in nycflights13::flights.

```{r}
looper("character", flights, class)
```


Compute the number of unique values in each column of iris.

```{r}
looper("numeric", iris, function(x) sum(table(unique(x))))
```

Generate 10 random normals for each of u = -10, 0, 10, and 100.

```{r}
ten_draws <- function(x) rnorm(10, mean = x)

map(c(-10, 0, 10, 100), ten_draws)
```


Think about the output, sequence, and body before you start writing the loop.

Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}

# Can be replaced with

str_c(out, collapse = "")

# ---

x <- sample(100)
sd <- 0
for (i in seq_along(x)) {
  sd <- sd + (x[i] - mean(x)) ^ 2
}
sd <- sqrt(sd / (length(x) - 1))

# can be replaced with 

sd(x)

# ---

x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]

for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}

# can be replaced with

cumsum(x)
```

Combine your function writing and for loop skills:

Write a for loop that prints() the lyrics to the children’s song "Alice the camel".

```{r}
song <-
  "Alice the camel has five humps.
Alice the camel has five humps.
Alice the camel has five humps.
So go, Alice, go.
"

split_song <- str_split(song, "\\.")[[1]]

for (i in split_song) {
  cat(i)
}
```


Convert the nursery rhyme "ten in the bed" to a function. Generalise it to any number of people in any sleeping structure.

```{r}

nurs <- function(x) {
  
  empty_list <- vector("list", x)

  empty_list[[1]] <- '\n There was 1 in the bed \n And the little one said, "Alone at last!"'

    
  for (i in seq_len(x)[-1]) {
    the_str <-
      paste0('\n There were ', i,' in the bed \n And the little one said, \n "Roll over! Roll over!" \n So they all rolled over and one fell out \n')
    
    empty_list[[i]] <- the_str
  }
  
  for (i in rev(empty_list)) cat(i, sep = "\n")
}

nurs(10)
```


Convert the song "99 bottles of beer on the wall" to a function. Generalise to any number of any vessel containing any liquid on any surface.


```{r}
beers <- function(x, drink, where) {
  
  for (i in c(rev(seq_len(x)), 0)) {
    if (i == 0) {
      cat(glue::glue("No more bottles of {drink} on the {where}, no more bottles of {drink}.\n Go to the store and buy some more, {x} bottles of {drink} on the {where}."))
    } else {
      cat(glue::glue("{i} {if (i == 1) 'bottle' else 'bottles'} of {drink} on the {where}, {i} {if (i == 1) 'bottle' else 'bottles'} of {drink}.\n Take one down and pass it around, {if (i == 1) 'no more' else i - 1} bottles of {drink} on the {where}. \n \n \n"))
    }
  }
}

beers(10, 'beer', 'wall')
beers(10, 'wine', 'floor')
```


It’s common to see for loops that don’t preallocate the output and instead increase the length of a vector at each step:

```{r, results = 'asis'}
x <- rerun(10000, sample(sample(10, 1)))

microbenchmark::microbenchmark(
  non_allocate = {
  output <- vector("integer", 0)
  for (i in seq_along(x)) {
    output <- c(output, length(x[[i]]))
  }
}
,
  allocate = {
  output <- vector("integer", 0)
  for (i in seq_along(x)) {
    output[i] <- length(x[[i]])
  }
})
```

How does this affect performance? Design and execute an experiment.

The first (growing the vector) is nearly 15 times slower than allocating!

## Exercises 21.3.5

Imagine you have a directory full of CSV files that you want to read in. You have their paths in a vector, files <- dir("data/", pattern = "\\.csv$", full.names = TRUE), and now want to read each one with read_csv(). Write the for loop that will load them into a single data frame.

```{r, eval = FALSE}
all_csv <- c("one.csv", "two.csv")

all_dfs <- vector("list", length(all_csv))

for (i in all_csv) {
  all_dfs[[i]] <- read_csv(all_csv[[i]])
}

bind_rows(all_dfs)
```


What happens if you use for (nm in names(x)) and x has no names? What if only some of the elements are named? What if the names are not unique?

```{r}

no_names <- 1:5
some_names <- c("one" = 1, 2, "three" = 3)
repeated_names <- c("one" = 1, "one" = 2, "three" = 3)

for (nm in names(no_names)) print(identity(nm)) # nothing happens!
for (nm in names(some_names)) print(identity(nm)) # the empty name get's filled with a ""
for (nm in names(repeated_names)) print(identity(nm)) # everything get's printed out
```


Write a function that prints the mean of each numeric column in a data frame, along with its name. For example, show_mean(iris) would print:

```{r, eval = FALSE}
show_mean(iris)
#> Sepal.Length: 5.84
#> Sepal.Width:  3.06
#> Petal.Length: 3.76
#> Petal.Width:  1.20

show_means <- function(x) {
  
  the_class <- vector("logical", length(x))

  for (i in seq_along(x)) the_class[[i]] <- is.numeric(x[[i]])
  
  x <- x[the_class]
  
  for (i in seq_along(x)) {
    cat(paste0(names(x)[i], ": ", round(mean(x[[i]]), 2)), fill = TRUE)
  }
}

show_means(iris)
show_means(mtcars)
```

(Extra challenge: what function did I use to make sure that the numbers lined up nicely, even though the variable names had different lengths?)

`cat` with `fill` set to `TRUE`

What does this code do? How does it work?

```{r, eval = FALSE}

trans <- list( 
  disp = function(x) x * 0.0163871,
  am = function(x) {
    factor(x, labels = c("auto", "manual"))
  }
)
for (var in names(trans)) {
  mtcars[[var]] <- trans[[var]](mtcars[[var]])
}
```

It converts `disp` and `am` by multiplying and then into a factor respectively. This is simply iterating over a list with functions, and applying in that same order to both variables.

## Exercises 21.4.1

Read the documentation for apply(). In the 2d case, what two for loops does it generalise?

It allows to loop over rows or columns and apply a function to that specific vector.

Adapt col_summary() so that it only applies to numeric columns You might want to start with an is_numeric() function that returns a logical vector that has a TRUE corresponding to each numeric column.

```{r}
col_summary <- function(x) {
  
  the_numeric <- vector("logical", length(x))
  for (i in seq_along(x)) the_numeric[[i]] <- is.numeric(x[[i]])
  
  x <- x[the_numeric]
  
  the_mean <- vector("numeric", length(x))
  for (i in seq_along(x)) the_mean[[i]] <- round(mean(x[[i]]), 2)
  
  the_mean
}

col_summary(iris)
```

## Exercises 21.5.3
Write code that uses one of the map functions to:

Compute the mean of every column in mtcars.

```{r}
map_dbl(mtcars, mean)
```


Determine the type of each column in nycflights13::flights.

```{r}
map(flights, class)
```


Compute the number of unique values in each column of iris.

```{r}
map(iris, ~ length(unique(.)))
```

Generate 10 random normals for each of u  = -10, 0, 10, and 100.

```{r}
map(c(-10, 0, 10, 100), rnorm, n = 10)
```


How can you create a single vector that for each column in a data frame indicates whether or not it’s a factor?

```{r}
map_lgl(iris, is.factor)
```


What happens when you use the map functions on vectors that aren’t lists? What does map(1:5, runif) do? Why?

```{r}
map(1:5, runif)
```
It iterates the same way. `map` functions take either 2d or 1d objects and iterate over (columns) or the 1d object.

What does map(-2:2, rnorm, n = 5) do? Why? What does map_dbl(-2:2, rnorm, n = 5) do? Why?

`map(-2:2, rnorm, n = 5)` runs `rnorm` with mean `-2`, then `-1`, ..., `2` producing a vector of 5 numbers for each call. This returns a list because there's no prefix. Conversely, `map_dbl(-2:2, rnorm, n = 5)` repeates the same but checks that the result is double `_dbl` and of length one. 

Rewrite map(x, function(df) lm(mpg ~ wt, data = df)) to eliminate the anonymous function.

```{r, eval = F}
map(x, function(df) lm(mpg ~ wt, data = df))

map(x, ~ lm(mpg ~ wt, data = .))
```

## Exercises 21.9.3

Implement your own version of every() using a for loop. Compare it with purrr::every(). What does purrr’s version do that your version doesn’t?

```{r}
my_every <- function(.x, .f, ...) {
  the_test <- vector("logical", length(.x))
  for (i in seq_along(.x)) the_test[[i]] <- .f(.x[[i]], ...)
  all(the_test, na.rm = TRUE)
}

my_every(mtcars, is.numeric)
every(mtcars, is.numeric)
```
`every` is probably much much faster and allows to supply one-sided formulas as functions. I don't understand clearly what `.p` is about in the docs and I can't get it work properly.

```{r}
every(1:10, ~ rep(TRUE, 9)) # should throw an error?
```

Create an enhanced col_sum() that applies a summary function to every numeric column in a data frame.

```{r}
col_sum <- function(x, f, ...) {
  x <- keep(x, is.numeric)
  map_dbl(x, f, ...)
}

col_sum(mtcars, mean, trim = 0.8)
col_sum(iris, median)
```



A possible base R equivalent of col_sum() is:

```{r}
col_sum3 <- function(df, f) {
  is_num <- sapply(df, is.numeric)
  df_num <- df[, is_num]

  sapply(df_num, f)
}
```

But it has a number of bugs as illustrated with the following inputs:

```{r, eval = FALSE}
df <- tibble(
  x = 1:3, 
  y = 3:1,
  z = c("a", "b", "c")
)
# OK
col_sum3(df, mean)
# Has problems: don't always return numeric vector
col_sum3(df[1:2], mean)
col_sum3(df[1], mean)
col_sum3(df[0], mean)
```

What causes the bugs?

`sapply` returns different things based on input. `sapply(df[0], is.numeric)` returns a named list rather than a `tibble`, whereas `sapply(df[1], is.numeric)` returns the correct output.