---
title: "R-4DS"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 3.6.1 Exercises

What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
```{r}
# Line chart
mpg %>%
  group_by(year) %>%
  summarise(m = mean(cty)) %>%
  ggplot(aes(year, m)) + 
  geom_line()

# Boxplot
ggplot(mpg, aes(class, hwy)) +
  geom_boxplot()

# Histogram
ggplot(mpg, aes(displ)) +
  geom_histogram(bins = 60)

# Area chart
huron <- data.frame(year = 1875:1972, level = as.vector(LakeHuron))
ggplot(huron, aes(year, level)) +
  geom_area()

```

Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

What does show.legend = FALSE do? What happens if you remove it?
Why do you think I used it earlier in the chapter?

It removes the legend.

What does the se argument to geom_smooth() do?
It removes the confidence intervals from the smoothed lines

Will these two graphs look different? Why/why not?
```{r, echo = F}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They'll be exactly the same.

Recreate the R code necessary to generate the following graphs.

```{r}
# 1st.
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(se = F)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth(aes(group = drv), se = F)

# 2nd.
ggplot(mpg, aes(displ, hwy, colour = drv)) +
  geom_smooth(se = F) +
  geom_point()

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(se = F)

# 3rd.
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = drv)) +
  geom_smooth(aes(linetype = drv), se = F)

# You can do this one by choosing a shape which has a border and simply colour
# the border with `colour` and the insides with `fill` (which is matched to drv).
# Then make the whole point a bit bigger with size
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(fill = drv), shape = 21, stroke = 2, colour = "white", size = 3)
```

## 3.7.1 Exercises

What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?

```{r, echo = F}
# Previous plot
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

`stat_summary` is associated with `geom_pointrange`.

```{r}
ggplot(diamonds) +
  geom_pointrange(aes(cut, depth, ymin = depth, ymax = depth))
```

What does geom_col() do? How is it different to geom_bar()?

`geom_col` leaves the data as it is. `geom_bar()` creates two variables (count and prop) and then graphs the count data on the y axis. With `geom_col` you can plot the values of any x variable against any y variable.

```{r}
# For example, plotting exactly x to y values.
aggregate.data.frame(diamonds$price, list(diamonds$cut), mean, na.rm = T) %>%
  print(.) %>%
  ggplot(aes(Group.1, x)) +
  geom_col()
```

Most geoms and stats come in pairs that are almost always used in concert. Read through the documentation and make a list of all the pairs. What do they have in common?

What variables does stat_smooth() compute? What parameters control its behaviour?

`stat_smooth()` computes the slope of the relationship x and y. It can compute different methods such as `lm`, `glm`, `lowess` among others. See method in `?stat_smooth`. The statistic can be controlled with the method argument.

In our proportion bar chart, we need to set group = 1. Why? In other words what is the problem with these two graphs?

```{r}
ggplot(data = diamonds)
  geom_bar(mapping = aes(x = cut, y = ..prop..))

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = color, y = ..prop..))

m <- ggplot(data = diamonds)
m + geom_bar(aes(cut, ..count../sum(..count..)))

diamonds %>%
  count(cut) %>%
  mutate(prop = n/sum(n)) %>%
  ggplot(aes(cut, prop)) + geom_bar(stat = "identity")

ggplot(diamonds, aes(cut)) + geom_bar(aes(y = ..count../sum(..count..)))

ggplot(diamonds, aes(cut)) + geom_bar(aes(y = ..prop.., group = 1))
```

## 3.8.1 Exercises

What is the problem with this plot? How could you improve it?

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```
Althought the two variables are continuous, the chance of being in a single point is very discrete and a lot of points overlap. We could fix it by adding jitter.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_jitter()
```

What parameters to geom_jitter() control the amount of jittering?
`width` and `height`

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_jitter(width = 5, height = 10)
```

Compare and contrast geom_jitter() with geom_count().
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_jitter()

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point() +
  geom_count()
```
`geom_count()` is another variant of `geom_point()` and controls the size of each dot based on the frequency of observations in a specifiy coordinate. It can help to contrast with `geom_jitter()` in understanding the data.

What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = displ)) + 
  geom_boxplot(aes(colour = drv))
```

## 3.9.1 Exercises

Turn a stacked bar chart into a pie chart using coord_polar().

```{r}
p <- ggplot(mpg, aes(class, fill = as.factor(cyl))) +
  geom_bar()
p

p + coord_polar()
```

What does labs() do? Read the documentation.
`labs()` allows you to control all the labels in the plot. For example:
```{r}
p + labs(title = "Hey, this is a title",
         x = "This is the X axis",
         y = "This is the Y axis",
         fill = "This is the fill")
```

What’s the difference between coord_quickmap() and coord_map()?
```{r}
nz <- map_data("nz")

nzmap <- ggplot(nz, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")
nzmap + coord_map()
nzmap + coord_quickmap()
```

`coord_quickmap()` is very similar to `coord_map()` but `coord_quickmap()` preserves straight lines in what should be a spherical plane. So, basically, the earth is shperical and `coord_map()` preserves that without plotting any straight lines. `coord_quickmap()` adds those lines.

What does the plot below tell you about the relationship between city and highway mpg? Why is coord_fixed() important? What does geom_abline() do?

```{r}
p <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline()
p
p + coord_fixed()
```

There is a positive correlation between the two. `coord_fixed()` makes sure there is no visual discrepancies and 
> ensures that the ranges of axes are equal to the specified ratio by adjusting the plot aspect ratio - Documentation of `coord_fixed()`.

Finally, `geom_abline()` plots the estiamted slope between the two variables.
