---
title: "Ch7"
output: html_notebook
---

```{r}
library(tidyverse)
```

7.3.4 Exercises

Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
diamonds %>%
  gather(key = dist, vals, x, y, z) %>%
  ggplot(aes(vals, colour = dist)) +
  geom_freqpoly(bins = 100)
```

One thing that is pretty obvious but perhaps hard to grasp at first is that the distribution of X and Y are pretty much the same. In fact, the same graph from above with `bins = 30` won't show you the X distribution because it overlaps perfectly. The correlation between the two is `cor(diamonds$x, diamonds$y)`.

If we rounded each mm to the closest number, value-pairing x and y yields `mean(with(diamonds, round(x, 0) == round(y, 0)))` of the values with the same number. So far, the length is directly proportional to the y value.

```{r}
diamonds %>%
  filter(y < 30) %>%
  select(x, y, z) %>%
  ggpairs()
```

Yet the relationship between x and y with z is almost flat, as expected. That is, after excluding 2 diamonds which had a unreasonable values.

Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
## TODO: Fix the Y and X axis to be able to specify the cutting point in the distribution.
source("http://peterhaschke.com/Code/multiplot.R")

graph <- map(seq(50, 1000, 100),
             ~ ggplot(diamonds, aes(x = price)) +
               geom_histogram(bins = .x) +
               labs(x = NULL, y = NULL) +
               scale_x_continuous(labels = NULL) +
               scale_y_continuous(labels = NULL))

multiplot(plotlist = graph, cols = 1)
```


How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?

Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?

