---
title: "Ch7"
output:
  html_document:
    df_print: paged
---

## 10.5 Exercises

How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame)

Visually you can tell by the printing method and how it stores the class of each variable underneath each variable. Or just get the `class()` of object and see whether it has `tbl_df` and `tbl`.

Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x # on a tibble this will throw a warning. Partial matching is not allowed
df[, "xyz"] # on a tibble this will be a data frame still
df[, c("abc", "xyz")] # This will be the same result in a tibble
```

The frustration is because in some situations a data frame will returns a different thing like in the last two previous lines of code. Tibble will return the same thing, providing consistency.

If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"
as_tibble(mtcars)[[var]]

# or

as_tibble(mtcars)[var]
```


Practice referring to non-syntactic names in the following data frame by:

```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

Extracting the variable called 1.

```{r}
annoying$`1`
```


Plotting a scatterplot of 1 vs 2.

```{r}
ggplot(annoying, aes(`1`, `2`)) +
  geom_point()
```

Creating a new column called 3 which is 2 divided by 1.

```{r}
annoying <-
  annoying %>%
  mutate(`3` = `2` / `1`)
```


Renaming the columns to one, two and three.

```{r}
annoying %>%
  rename(one = `1`,
         two = `2`,
         three = `3`)
```

What does `tibble::enframe()` do? When might you use it?

It turns named vectors or list to two-column data frames.

It's different from `as_tibble()` for lists because it creates a stacked data frame rather than a widy one. It all dependes on your data.
```{r}
lst <- list(female = 1, male = 2)

as_tibble(lst)
```
rather than

```{r}
enframe(lst) %>% unnest()
```


What option controls how many additional column names are printed at the footer of a tibble?

`options(tibble.width = Inf)` for all columns to be printed.