---
title: "Ch7"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```


## Exercises 11.3.5

What function would you use to read a file where fiels are separated with "|"?

You use `read_delim` and specify "|" in the `delim` argument.

Apart from `file`, `skip` and `comment`, what other arguments do `read_csv()` and `read_tsv()` have in common?

All arguments! But that's logical because they both use `read_delim` as the function doing the work. Both functions just call `read_delim` with a set of predefine options for the `csv` and `tsv` formats using the `tokenize_*` functions. The `tokenize_*` functions simply return a list with the charachteristics of each format.

What are the most important arguments to `read_fwf()`?

The most important argument is `col_positions` because that's how determine the width at which each column is separated. You can determine the width with the `fwf_*` helper functions.

What arguments do you need to specify to read the following text into a data frame?

```{r}
read_csv("x,y\n1,'a,b'", quote = "'")
read_delim("x,y\n1,'a,b'", delim = ",",  quote = "'")
```

Identify what is wrong with each of the following inline CSV files. What happens when you run the code?

```{r}
read_csv("a,b\n1,2,3\n4,5,6") # more rows then column names
read_csv("a,b\n1,2,3\n4,5,6", skip = 1, col_names = letters[1:3]) # fixed

read_csv("a,b,c\n1,2\n1,2,3,4") # second row has only two values but the remaining lines have 3

read_csv("a,b\n\"1") # the second row is actually: ", 1. but it uses \" so it's a literal " and a comma is missing
read_csv('a,b\n\",1', quote = "'") # it should be something like this I think

read_csv("a,b\n1,2\na,b") # nothing wrong with this one. Maybe the column classes because a and b are column names errors
read_csv("a,b\n1,2\na,b", n_max = 1) # this is the correct format.

read_csv("a;b\n1;3") # this is ; deliminted
read_csv2("a;b\n1;3")
```


# Exercise 11.3.5

What are the most important arguments to `locale()`?

This is a bitr tricky because they're all imporant. `encoding` is usually one that brings about a lot of problems if you're working with international data. However, all others are also important except for `asciify` which is hardly used.

What happens if you try and set `decimal_mark` and `grouping_mark` to the same character? What happens to the default value of `grouping_mark` when you set `decimal_mark` to ","? And viceversa?

If you set both to the same character it throws an error. Why? It makes sense. How can you distinguish cents from thousands here? `123,456,789`? Is this 123 million or 123 thousand?

```{r}
locale(decimal_mark = ".")
locale(decimal_mark = ",")
```

`.` and `,` are the defaults. If you change one the other gets assigned.

I didn't discuss the `date_format` and `time_format` options to `locale()`. What do they do? Construct an example that shows when they might be useful?

They set the default date/time formats, which are based on the `ISO8601` format of yyyy-mm-dd hh-mm-ss. You can override that default by specifying the `locale` argument with new defaults.

For date:
```{r}
read_csv("a\n05-02-00")
read_csv("a\n05-02-00", locale = locale(date_format = "%d-%m-%y"))
```

For time:
```{r}
read_csv("a\n02-00-08 am")
read_csv("a\n02-00-08 am", locale = locale(time_format = "%M-%S-%I %p"))
```


If you live outside the U.S, create a new `locale` object that encapsulates the type of files you read more often.

```{r}
locale(date_names = "es",
       date_format = "%Y/%m/%d",
       time_format = "%H/%M/%S",
       grouping_mark = ".")
```

What's the difference between `read_csv()` and `read_csv2()`? They performe exactly the same operation but the first reads comma-delimited files and the second one read semi-colon delimited files.

What are the most common encodings in Europe? and in Asia?

No internet right now!

Generate the correct format string to parse each of the following dates and times:

```{r}  
d1 <- "January 1, 2010"
parse_date(d1, format = "%B %d, %Y")

d2 <- "2015-Mar-07"
parse_date(d2, "%Y-%b-%d")

d3 <- "06-Jun-2017"
parse_date(d3, "%d-%b-%Y")

d4 <- c("August 19 (2015)", "July 1 (2015)")
parse_date(d4, "%B %d (%Y)")

d5 <- "12/30/14"
parse_date(d5, "%m/%d/%y")

t1 <- "1705"
parse_time(t1, "%H%M")

t2 <- "11:15:10.12 PM"
parse_time(t2, "%I:%M:%OS %p")
```

