---
title: "Ch9"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```


# Exercise 12.2.1

Using prose describe how each of variables and observations are organized in the the tables.

Table 1:
```{r}
table1
```

All columns are it's own variables because for example `year` contains only years and `country` contains only countries. For example, if we got a data set where we have `country` and then two columns for `males` and `females` that would be untidy. At least in principle. Because both columns should be a column called `gender`. This is a bit tricky because nothing says that you can't use the gender columns that way. But for easy use in R it's usually better to work with `tidy` data. But who says that `cases` and `populations` are not the same thing? We could argue that they should be in the same column as it is now.

```{r}
table2
```

In principle, this is also not a `tidy` dataset, although that's debatable. If we had something like different sicknesses in the type column then this would be tidy. But `type` in this case should be different columns because they measure different things. This dataset is organized in a way that years are nested within countries and then each type is instead within years.

```{r}
table3
```
This is clearly non-tidy because we can't work with values such as the `rate` column. If this column would be the result of the operation, then this would be a tidy dataset.

```{r}
table4a
```
This is clearly not a tidy dataset because years, which are the same thing, are in different columns. This is the same case as the gender example I outlined above. These two columns should be *one* variable and the values should be a separate column. This dataset is only for the type `cases`

```{r}
table4b
```

Same as above, but this is for `population`. 

Compute the `rate` for `table2`, and `table4a` and `table4b`.

For table 2
```{r}
fyear_cases <-
  table2 %>%
  filter(year == 1999,
         type == "cases")

fyear_pop <-
  table2 %>%
  filter(year == 1999,
         type == "population")

rate_99 <-
  bind_cols(fyear_cases, fyear_pop) %>%
  mutate(rate = count/count1) %>%
  select(-ends_with("1"), -type)


fyear_cases <-
  table2 %>%
  filter(year == 2000,
         type == "cases")

fyear_pop <-
  table2 %>%
  filter(year == 2000,
         type == "population")

rate_00 <-
  bind_cols(fyear_cases, fyear_pop) %>%
  mutate(rate = count/count1) %>%
  select(-ends_with("1"), -type)

bind_rows(rate_99, rate_00)
```

For table 4a and 4b

```{r}
table4a %>%
  mutate(rate_99 = `1999` / table4b$`1999`,
         rate_00 = `2000` / table4b$`2000`)
```

In a way, it's more intuitive to work with `table2` because we use the filtering techniques to understand the operations. But with `table4a` is more succient. However, if we wanted to turn the `table4a/4b` result to a tidier version, it would be more tedious (withut `gather` and such functions.)

Recreate the plot showing change in cases over time using `table2` instead of `table1`. What do you need to do first?

```{r}
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count, group = country, colour = country)) +
    geom_point() +
    geom_line()
```

## 12.3.3 Exercises

Why are gather() and spread() not perfectly symmetrical?

```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```
Both spread() and gather() have a convert argument. What does it do?

Because the `key` variable is actually the column names, and is thus moved as character column. It would be unwise for gather to treat column names as numerics, logicals, or something else. However, you can find a workaround by specifying `convert = TRUE` which will try to convert the `key` columns to it's correct class.

Why does this code fail?

```{r}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
```

Because `gather` can't find the columns names. You can't name columns w/ numbers in R without quoting them with tick marks.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```


Why does spreading this tibble fail? How could you add a new column to fix the problem?

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

Because Phillip Woods has two values of age. Think about it.. Phillip woods then would have TWO columns of age. That doesn't make sense! We need to add a unique column id specifying the third or first age as a unique person.

```{r}
people %>%
  mutate(unique_id = c(1, 2, 2, 3, 3)) %>%
  select(unique_id, everything()) %>%
  spread(key, value)
```

Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
```

The main objective of analysis here is whether pregnant or not (bc males can not be pregnant), so I would go for `gather`ing the gender column rather than spreading the pregnant column.

```{r}
preg %>%
  gather(gender, values, -pregnant)

# the other way around:
preg %>%
  gather(gender, values, -pregnant) %>%
  spread(pregnant, values)
```

## 12.4.3 Exercises

What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), fill = "warn")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
```

It's simple. x has vectors with 3 and 4 characters but we specify 3 columns. `fill` has three values: 
`warn`, `right` and `left`. Here I specify a fourth column to place the extra letter. The first fills the missing values with the extra character using the right most match. `left` does the same thing but without a warning. and left places the extra character empty in the first column

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three", "four"))

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three", "four"), fill = "right")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three", "four"), fill = "left")
```
I've deleted the fourth column to see how this works. `extra` on the other hand, deals with either droping or merging the extra characters. `warn` drops the extra character and emits a warning messge.
`drop` does the same thing but without a warning and `merge` merges the extra character to it's closest end. No aparent option to `merge` with the first column rather than the last.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "warn")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")
```


Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?

Because `unite` and `separate` receive columns and create new ones, `remove` allows you to remove the original columns that you unite/separate on. You might want to leave them as they are if you're checking whether the transformation was done correctly.

Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?

Because you can separate differently. Examples below:

```{r}
df_sep <- data.frame(x = c(NA, "a-b", "a-d", "b-c", "d-e"))
df_extract <- data.frame(x = c(NA, "ap.b", "aa/d", "b.c", "d-ee"))

# This is easy with separate
df_sep %>% separate(x, c("new", "old"), sep = 1)
df_sep %>% separate(x, c("new", "old"), sep = "-")

# Here we can define 2 or more groups to separate the more complex string
df_extract %>% extract(x, c("new", "old"), regex = "(.*)[:punct:](.*)")
```
However, I don't understand it completely because I think I could do the same as above with `separate` by just providing a regular expression.
