---
title: "Ch9"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
```


# Exercise 12.2.1

Using prose describe how each of variables and observations are organized in the the tables.

Table 1:
```{r}
table1
```

All columns are it's own variables because for example `year` contains only years and `country` contains only countries. For example, if we got a data set where we have `country` and then two columns for `males` and `females` that would be untidy. At least in principle. Because both columns should be a column called `gender`. This is a bit tricky because nothing says that you can't use the gender columns that way. But for easy use in R it's usually better to work with `tidy` data. But who says that `cases` and `populations` are not the same thing? We could argue that they should be in the same column as it is now.

```{r}
table2
```

In principle, this is also not a `tidy` dataset, although that's debatable. If we had something like different sicknesses in the type column then this would be tidy. But `type` in this case should be different columns because they measure different things. This dataset is organized in a way that years are nested within countries and then each type is instead within years.

```{r}
table3
```
This is clearly non-tidy because we can't work with values such as the `rate` column. If this column would be the result of the operation, then this would be a tidy dataset.

```{r}
table4a
```
This is clearly not a tidy dataset because years, which are the same thing, are in different columns. This is the same case as the gender example I outlined above. These two columns should be *one* variable and the values should be a separate column. This dataset is only for the type `cases`

```{r}
table4b
```

Same as above, but this is for `population`. 

Compute the `rate` for `table2`, and `table4a` and `table4b`.

```{r}
fyear <-
  table2 %>%
  filter(year == 1999)

fyear %>%
  filter(type == "cases")
```


Exercise 3
```{r}
table2 %>%
  filter(type == "cases") %>%
  ggplot(aes(year, count, group = country, colour = country)) +
    geom_point() +
    geom_line()
```

