---
title: "R-4DS"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 3.2.4 Exercises

Run ggplot(data = mpg) what do you see?
```{r}
ggplot(data = mpg)
```

How many rows are in mtcars? How many columns?
```{r}
dim(mpg)
```
What does the drv variable describe? Read the help for ?mpg to find out.

Whether the car is front wheel drive or not.
f = front-wheel drive, r = rear wheel drive, 4 = 4wd

Make a scatterplot of hwy vs cyl.
```{r}
ggplot(mpg) + geom_point(aes(hwy, cyl))
```
What happens if you make a scatterplot of class vs drv. Why is the plot not useful?
```{r}
ggplot(mpg) + geom_point(aes(class, drv))
```

Because both variables are categorical.

## 3.3.1 Exercises

What’s gone wrong with this code? Why are the points not blue?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?

### Categorical
- Model
- cyl
- Manufacturer
- trans
- drv
- fl
- class

### Continuous
- displ
- year
- cty
- hwy

Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = cty))
# ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, shape = cty)) This creates an error
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```

What happens if you map the same variable to multiple aesthetics?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = cty, size = cty))
```

What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)

Stroke controls the width of the border of certain shapes. Those shapes which have borders are the only ones that stroke can alter.

What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, colour = displ < 5))
```

## 3.5.1 Exercises

What happens if you facet on a continuous variable?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ cty)
```

It plots it anyway

What do the empty cells in plot with facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

It means that for cyl == 5 and drv == 4 | drv == r there are no cars. Also, there are no cars with 7 cyl.

What plots does the following code make? What does . do?

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

The dot controls whether the facetting will be done row or column wise. For example `facet_grid(drv ~ .)` will use drv as rows while `facet_grid(. ~ drv)` will use it as columns. `facet_grid(~ drv)` will do the same as the column wise facetting but `facet_grid(drv ~)` won't because a formula object needs to have something after the `~`.

Take the first faceted plot in this section:

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?

I think facetting is better when you want to pay particular attention to particular facets alone (naturally) while using the color aesthetic is better to discriminate which points are located where.

Read ?facet_wrap. What does nrow do? What does ncol do? What other options control the layout of the individual panels? Why doesn’t facet_grid() have nrow and ncol variables?

`nrow` controls the number of rows for the total number of facets whereas `ncol` controls the number of columns. Other options can control interesting parameters. For example, scales can control whether each plot has its own y axis with `scales = "free"`, as in allow the axes to be free. The function also has the labeller option to change the names of each facet and other options like `strip.position` for the position of the facets labels. Read `?facet_wrap` for more options.

`facet_grid` doesn't have the option to specify rows or columns because it calculate automatically the grid. So the multiplication of the number of distinct values in the variables in the formula.


When using facet_grid() you should usually put the variable with more unique levels in the columns. Why?

Because otherwise the graph is going to be too long and you won't understand anything. This graph is a good example:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(~ model)
```


